---
title: "Parallel Trends Graph"
author: "DANIEL PÉREZ"
date: "2024-05-20"
output:
  html_document: default
  pdf_document: default
---
```{r}
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(geosphere)
library(climaemet)
library(ggplot2)
```

I am repeating the same process as in script 1, where I create the dataframe. I am running everything from scratch again because after creating the dataframe and conducting the causal analysis, I decided to add data from 2015 and 2016, which allow me to improve the parallel trends graphs. Therefore, to avoid any issues with the work I have done, I have decided to repeat the process from the beginning to ensure that the scripts are still working correctly.


MADRID CITY CENTER DATA
```{r}
datos_lista <- list()

for (year in 2015:2023) {
 
  file_name <- paste0(year, ".csv")
  datos <- read_delim(file_name, delim = ";", escape_double = FALSE, trim_ws = TRUE)
  datos_lista[[year - 2014]] <- datos
}
```

```{r}
df_final <- do.call(rbind, datos_lista)
```


```{r}
daily_pollution <- df_final |> 
    rename(town_code = MUNICIPIO,
           station_code = ESTACION,
           magnitude = MAGNITUD,
           year = ANO,
           month = MES,
           provincia = PROVINCIA,
           punto_muestreo = PUNTO_MUESTREO) |> 
    select(-matches("^V"))

daily_pollution <- daily_pollution |> 
  filter(magnitude == "8") |> 
    select(-magnitude)
```


```{r}
daily_pollution <- daily_pollution |> 
  pivot_longer(cols = starts_with("D"),
               names_to = "dia",
               values_to = "no2_pollution")
```


```{r}
daily_pollution_center <- daily_pollution

for (i in 1:nrow(daily_pollution_center)) {
  if (substr(daily_pollution_center$dia[i], 1, 1) == "D") {
    daily_pollution_center$dia[i] <- substr(daily_pollution_center$dia[i], 2, nchar(daily_pollution_center$dia[i]))
  }
}
```

```{r}
fecha_str <- paste(daily_pollution_center$year, daily_pollution_center$month, daily_pollution_center$dia, sep = "-")

# Convertir la cadena de fecha a un objeto de fecha
daily_pollution_center$fecha <- as.Date(fecha_str, format = "%Y-%m-%d")

```
Calculate the average NO2 Pollution each year
```{r}
daily_pollution_center <- daily_pollution_center |> 
  mutate(
    no2_pollution = as.numeric(no2_pollution),
    fecha = as.factor(fecha))

mean_no2_pollution_per_year <- daily_pollution_center |> 
  group_by(year) |> 
  summarize(mean_no2_pollution = mean(no2_pollution, na.rm = TRUE))


mean_no2_pollution_per_year
  
```

DATA FROM THE COMMUNITY OF MADRID, NOT JUST THE CITY CENTER

```{r}
datos_lista_OUT <- list()
for (year in 2015:2023) { 
  file_name <- paste0(year, "_out.csv")  
  datos_OUT <- read_delim(file_name, delim = ";", escape_double = FALSE, trim_ws = TRUE)
  datos_lista_OUT[[year - 2014]] <- datos_OUT
}
```

Verify that all dataframes have the same columns.

```{r}
columnas_lista_OUT <- lapply(datos_lista_OUT, names)
columnas_lista_OUT

```

We see that the 2015 data has a column called ‘v24,,,,,,,’ when it should be just V24. So we correct it and put it all together in the same dataframe.

```{r}
datos_lista_OUT <- list()

for (year in 2015:2023) {
  file_name <- paste0(year, "_out.csv")
  datos_OUT <- read_delim(file_name, delim = ";", escape_double = FALSE, trim_ws = TRUE)
  # Correct problem
  if (year == 2015) { 
    names(datos_OUT)[names(datos_OUT) == "v24,,,,,,,,,,,,,,,,,,,,,,,,"] <- "v24"
  }
  
  datos_lista_OUT[[year - 2014]] <- datos_OUT
}

#Combine dataframes
df_final_out <- do.call(rbind, datos_lista_OUT)

```

```{r}
daily_pollution_out <- df_final_out |> 
    rename(town_code = municipio,
           station_code = estacion,
           magnitude = magnitud,
           year = ano,
           month = mes
          ) |> 
    select(-matches("^V"))


daily_pollution_out <- daily_pollution_out |> 
  filter(magnitude == "8") |> 
    select(-magnitude)
```


```{r}
daily_pollution_out <- mutate(daily_pollution_out,
                               h01 = as.numeric(gsub(",", ".", h01)),
                               h02 = as.numeric(gsub(",", ".", h02)),
                               h03 = as.numeric(gsub(",", ".", h03)),
                               h04 = as.numeric(gsub(",", ".", h04)),
                               h05 = as.numeric(gsub(",", ".", h05)),
                               h06 = as.numeric(gsub(",", ".", h06)),
                               h07 = as.numeric(gsub(",", ".", h07)),
                               h08 = as.numeric(gsub(",", ".", h08)),
                               h09 = as.numeric(gsub(",", ".", h09)),
                               h10 = as.numeric(gsub(",", ".", h10)),
                               h11 = as.numeric(gsub(",", ".", h11)),
                               h12 = as.numeric(gsub(",", ".", h12)),
                               h13 = as.numeric(gsub(",", ".", h13)),
                               h14 = as.numeric(gsub(",", ".", h14)),
                               h15 = as.numeric(gsub(",", ".", h15)),
                               h16 = as.numeric(gsub(",", ".", h16)),
                               h17 = as.numeric(gsub(",", ".", h17)),
                               h18 = as.numeric(gsub(",", ".", h18)),
                               h19 = as.numeric(gsub(",", ".", h19)),
                               h20 = as.numeric(gsub(",", ".", h20)),
                               h21 = as.numeric(gsub(",", ".", h21)),
                               h22 = as.numeric(gsub(",", ".", h22)),
                               h23 = as.numeric(gsub(",", ".", h23)),
                               h24 = as.numeric(gsub(",", ".", h24)))


daily_pollution_out <- mutate(daily_pollution_out,
                               no2_pollution = rowMeans(select(daily_pollution_out, starts_with("h")), na.rm = TRUE)) 

daily_pollution_out <- mutate(daily_pollution_out,
                               no2_pollution = round(no2_pollution))


daily_pollution_out <- daily_pollution_out %>%
                         select(-starts_with("h"))

```

```{r}
fecha_str <- paste(daily_pollution_out$year, daily_pollution_out$month, daily_pollution_out$dia, sep = "-")

# Convertir la cadena de fecha a un objeto de fecha
daily_pollution_out$fecha <- as.Date(fecha_str, format = "%Y-%m-%d")
```

Average no2 pollution levels outside Madrid´s Municipality
```{r}
daily_pollution_out <- daily_pollution_out |> 
  mutate(
    no2_pollution = as.numeric(no2_pollution),
    fecha = as.factor(fecha))

mean_no2_pollution_per_year <- daily_pollution_out %>%
  group_by(year) %>%
  summarize(mean_no2_pollution = mean(no2_pollution, na.rm = TRUE))

# Mostrar el resultado
mean_no2_pollution_per_year
```

```{r}
daily_pollution_center$city_center <- 1
daily_pollution_out$city_center <- 0
df_pollution <- rbind(daily_pollution_center, daily_pollution_out)
unique_values2 <- unique(df_pollution$punto_muestreo)
print(unique_values2)
```
```{r}

df_pollution$no2_pollution <- as.numeric(df_pollution$no2_pollution)

df_pollution <- df_pollution[!is.na(df_pollution$fecha), ]

df_pollution <- df_pollution |> 
  mutate(punto_muestreo = str_replace(punto_muestreo, "_[^_]+_[^_]+$", ""))

#write.csv(df_pollution, file = "df_pollution.csv", row.names = FALSE)

```

INFORMACIÓN DE ESTACIONES CENTRO CIUDAD

```{r}
city_stations <- read_delim("informacion_estaciones_red_calidad_aire.csv", delim=";")

city_stations <- city_stations |> 
   select(
     punto_muestreo = CODIGO,
     station_code = CODIGO_CORTO,
     station_name = ESTACION,
     address = DIRECCION,
     station_type = NOM_TIPO,
     longitude = LONGITUD,
     latitude = LATITUD) %>% 
    mutate(station_code = as.character(station_code))


city_stations <- city_stations %>%
  mutate(
    station_type = str_replace_all(station_type, "tráfico", "trafico"),
    station_type = str_to_title(station_type),
    station_type = str_replace_all(station_type, "\\bfondo\\b", "Fondo"),
    station_type = str_replace_all(station_type, "\\bindustrial\\b", "Industrial"),
    station_type = str_replace_all(station_type, "\\btrafico\\b", "Trafico")
  )
```

INFORMACIÓN ESTACIONES FUERA DEL CENTRO DE MADRID

```{r}
com_stations <- read_delim("com_stations.csv", delim=";")
com_stations <- com_stations |> 
  mutate(
    punto_muestreo = estacion_codigo,
    station_code = str_sub(estacion_codigo, -2, -1),
    station_code = if_else(str_starts(station_code, "0"),
                           str_remove(station_code, "^0"),
                           station_code))

  com_stations <- com_stations |> 
  select(station_name = estacion_municipio,
          punto_muestreo,
          station_code,
          station_type = estacion_tipo_estacion,
          area_type = estacion_tipo_area,
          address = estacion_direccion_postal,
          longitude = estacion_coord_longitud,
          latitude = estacion_coord_latitud) 
  
com_stations <- com_stations %>%
  mutate(station_type = paste(area_type, station_type, sep = " ")) |> 
  select(-area_type)


# Fix the latitude and longitude columns using regular expressions
com_stations<- com_stations %>%
  # Remove quotes at the beginning and end of the longitude column
  mutate(longitude = str_remove(longitude, "^\"|\",\\s*\"$")) %>%
  # Replace commas with periods in the longitude column and convert to numeric
  mutate(longitude = as.numeric(gsub(",", ".", longitude))) %>% 
  # Remove quotes and commas from the end of the latitude column
  mutate(latitude = str_remove(latitude, "\"?,?$")) %>%
  # Replace commas with periods in the latitude column and convert to numeric
  mutate(latitude = as.numeric(gsub(",", ".", latitude))) 

# Remove any sequence of two commas AND quotation marks from the address column
com_stations <- com_stations%>% 
  # Remove quotes from the address column
  mutate(address = str_remove(address, '["\']')) %>% 
  # Remove any sequence of two commas from the address column
  mutate(address = str_remove(address, ",{2}"))
```


```{r}
unique_values1<- unique(daily_pollution_center$punto_muestreo)

unique_values3 <- unique(city_stations$punto_muestreo)

print(unique_values1)
print(unique_values3)

```


```{r}
unique_values2 <- unique(daily_pollution_out$punto_muestreo)
unique_values4 <- unique(com_stations$punto_muestreo)
print(unique_values2) 
print(unique_values4)
```
```{r}
# Lista de valores a eliminar
valores_a_eliminar <- c("28134002", "28115003", "28106001", "28127004")

# Filtrar el dataframe para eliminar las observaciones con esos valores en punto_muestreo
df_pollution <- df_pollution |> 
  filter(!punto_muestreo %in% valores_a_eliminar)

# Verificar los cambios
print(df_pollution)
```

Juntar ambos dataframes con informacion sobre las estaciones

```{r}
estaciones <- bind_rows(city_stations, com_stations)
```

```{r}
estaciones <- estaciones |> 
  mutate(punto_muestreo = as.character(punto_muestreo))

df_combined <- df_pollution |> 
  left_join(estaciones, by = "punto_muestreo")

```


```{r}
df_combined <- df_combined |>  drop_na(no2_pollution)

df_combined<- df_combined |> 
  filter(as.Date(fecha) <= as.Date("2021-06-30"))

colSums(is.na(df_combined))

```
```{r}
yearly_averages <- df_combined |> 
  group_by(year, city_center) |> 
  summarise(avg_yearly_pollution = mean(no2_pollution, na.rm = TRUE)) %>%
  ungroup()

yearly_averages <- yearly_averages |> 
  mutate(year = as.numeric(as.character(year)))

write.csv(yearly_averages, "yearly_averages.csv", row.names = FALSE)

ggplot(yearly_averages, aes(x = year, y = avg_yearly_pollution, color = as.factor(city_center), group = city_center)) + 
  geom_line() + 
  geom_point() +
  labs(x = "Year", 
       y = expression("Yearly Pollution Average (NO2)"), 
       title = expression("Examining Parallel Trends: Evolution of Yearly Average Pollution (NO2) Levels (2015 - 2021)"),
       color = "City center") + 
  geom_vline(xintercept = 2018.91, 
             linetype = "dashed", 
             color = "black") +
  annotate("text", x = 2019, y = max(yearly_averages$avg_yearly_pollution, na.rm = TRUE) * 0.95, 
           label = "Policy Implementation", vjust = -0.5, hjust = 0, size = 3) +
  theme_light() +
  theme(plot.title = element_text(size = 11),
        axis.text.x = element_text(angle = 45, hjust = 1))


```

```{r, echo=FALSE}
df_combined <- df_combined |> 
  mutate(zone = case_when(
    station_name %in% c("Escuelas Aguirre", "Castellana", "Plaza Castilla", 
                        "Ramón y Cajal", "Cuatro Caminos", "Plaza de España", 
                        "Barrio del Pilar", "Plaza del Carmen", "Méndez Álvaro", 
                        "Parque del Retiro") ~ "Zone 1",
    station_name %in% c("Plaza Elíptica", "Farolillo", "Villaverde", "Moratalaz", "Vallecas", "Ensanche de Vallecas") ~ "Zone 2",
    station_name %in% c("Arturo Soria", "Sanchinarro", "Urb. Embajada", 
                        "Barajas Pueblo", "Tres Olivos", "Juan Carlos I") ~ "Zone 3",
    station_name %in% c("El Pardo", "Casa de Campo") ~ "Zone 4",
    station_name %in% c("Valdemoro", "Arganda del Rey", "Collado Villalba", "Algete", "Colmenar Viejo", "Alcala de Henares") ~ "Control",
    TRUE ~ "Excluded"
  ))

```

```{r}

yearly_averages <- df_combined |> 
  group_by(year, zone) |> 
  summarise(avg_yearly_pollution = mean(no2_pollution, na.rm = TRUE)) %>%
  ungroup()

yearly_averages <- yearly_averages |> 
  mutate(year = as.numeric(as.character(year)))


zone_colors <- c("Zone 1" = "lightblue", "Zone 2" = "orange", "Zone 3" = "purple", 
                  "Zone 4" = "yellow", "Control" = "red", "Excluded" = "grey")


ggplot(yearly_averages, aes(x = year, y = avg_yearly_pollution, color = zone)) + 
  geom_line() + 
  labs(x = "Year", 
       y = expression("Yearly Pollution Averages"), 
       title = expression("Examining Parallel Trends: Evolution of Yearly Average Pollution Levels by Zone (2015 - 2021)"),
       color = "Zone") + 
  scale_color_manual(values = zone_colors) +
  geom_vline(xintercept =  2018.91, 
             linetype = "dashed", 
             color = "black") +
  annotate("text", x = 2019, y = max(yearly_averages$avg_yearly_pollution, na.rm = TRUE) * 0.95, 
           label = "Policy Implementation", vjust = -0.5, hjust = 0, size = 3) +
  theme_light() +
  theme(plot.title = element_text(size = 11))



```

