---
title: "TFM"
author: "DANIEL PÉREZ"
date: "2024-04-05"
output: html_document
---


```{r}
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(geosphere)
library(climaemet)
library(mice)
```

MADRID CITY CENTER DATA
```{r}
# Initialize a list to store the data from all files
datos_lista <- list()

# Loop to read each file and store the data in the list
for (year in 2017:2023) {
  # Generate the file name for the current year
  file_name <- paste0(year, ".csv")
  
  # Read the CSV file with read_delim
  datos <- read_delim(file_name, delim = ";", escape_double = FALSE, trim_ws = TRUE)
  
  # Store the data in the list
  datos_lista[[year - 2015]] <- datos
}
```

```{r}
df_final <- do.call(rbind, datos_lista)
```


```{r}
daily_pollution <- df_final |> 
    rename(town_code = MUNICIPIO,
           station_code = ESTACION,
           magnitude = MAGNITUD,
           year = ANO,
           month = MES,
           provincia = PROVINCIA,
           punto_muestreo = PUNTO_MUESTREO) |> 
    select(-matches("^V")) #remove columns from the daily_pollution dataframe that start with the letter "V"

#Filter for magnitude 8, which is NO2.
daily_pollution <- daily_pollution |> 
  filter(magnitude == "8") |> 
    select(-magnitude)
```


```{r}
daily_pollution <- daily_pollution |> 
  pivot_longer(cols = starts_with("D"),
               names_to = "dia",
               values_to = "air_quality")
```


```{r}
# Copy of the dataframe to avoid modifying the original
daily_pollution_center <- daily_pollution

# Remove the "D" from each observation in the "dia" column if it is present
for (i in 1:nrow(daily_pollution_center)) {
  if (substr(daily_pollution_center$dia[i], 1, 1) == "D") {
    daily_pollution_center$dia[i] <- substr(daily_pollution_center$dia[i], 2, nchar(daily_pollution_center$dia[i]))
  }
}
```

```{r}
#Create date column
fecha_str <- paste(daily_pollution_center$year, daily_pollution_center$month, daily_pollution_center$dia, sep = "-")

#Convert fecha to a date object
daily_pollution_center$fecha <- as.Date(fecha_str, format = "%Y-%m-%d")

```


DATA FROM THE COMMUNITY OF MADRID, NOT JUST THE CITY CENTER


```{r}
#Same process as for Madrid Municipality data
datos_lista_OUT <- list()

for (year in 2017:2023) { 
  file_name <- paste0(year, "_out.csv")  
  datos_OUT <- read_delim(file_name, delim = ";", escape_double = FALSE, trim_ws = TRUE)
  datos_lista_OUT[[year - 2015]] <- datos_OUT
}
```

```{r}
df_final_out <- do.call(rbind, datos_lista_OUT)
```

```{r}
daily_pollution_out <- df_final_out |> 
    rename(town_code = municipio,
           station_code = estacion,
           magnitude = magnitud,
           year = ano,
           month = mes
          ) |> 
    select(-matches("^V"))


daily_pollution_out <- daily_pollution_out |> 
  filter(magnitude == "8") |> 
    select(-magnitude)
```


```{r}
#Convert the hours variable from character type to numeric type

daily_pollution_out <- mutate(daily_pollution_out,
                               h01 = as.numeric(gsub(",", ".", h01)),
                               h02 = as.numeric(gsub(",", ".", h02)),
                               h03 = as.numeric(gsub(",", ".", h03)),
                               h04 = as.numeric(gsub(",", ".", h04)),
                               h05 = as.numeric(gsub(",", ".", h05)),
                               h06 = as.numeric(gsub(",", ".", h06)),
                               h07 = as.numeric(gsub(",", ".", h07)),
                               h08 = as.numeric(gsub(",", ".", h08)),
                               h09 = as.numeric(gsub(",", ".", h09)),
                               h10 = as.numeric(gsub(",", ".", h10)),
                               h11 = as.numeric(gsub(",", ".", h11)),
                               h12 = as.numeric(gsub(",", ".", h12)),
                               h13 = as.numeric(gsub(",", ".", h13)),
                               h14 = as.numeric(gsub(",", ".", h14)),
                               h15 = as.numeric(gsub(",", ".", h15)),
                               h16 = as.numeric(gsub(",", ".", h16)),
                               h17 = as.numeric(gsub(",", ".", h17)),
                               h18 = as.numeric(gsub(",", ".", h18)),
                               h19 = as.numeric(gsub(",", ".", h19)),
                               h20 = as.numeric(gsub(",", ".", h20)),
                               h21 = as.numeric(gsub(",", ".", h21)),
                               h22 = as.numeric(gsub(",", ".", h22)),
                               h23 = as.numeric(gsub(",", ".", h23)),
                               h24 = as.numeric(gsub(",", ".", h24)))


#As we have hourly data and from the city center we have daily data, I take the daily average

daily_pollution_out <- mutate(daily_pollution_out,
                               air_quality = rowMeans(select(daily_pollution_out, starts_with("h")), na.rm = TRUE)) 

daily_pollution_out <- mutate(daily_pollution_out,
                               air_quality = round(air_quality))

#As hours stars with h, I remove it
daily_pollution_out <- daily_pollution_out |> 
                         select(-starts_with("h"))

```


```{r}
fecha_str <- paste(daily_pollution_out$year, daily_pollution_out$month, daily_pollution_out$dia, sep = "-")
daily_pollution_out$fecha <- as.Date(fecha_str, format = "%Y-%m-%d")
```

```{r}
#Create dummy variable to indicate if it is city center or peripherial areas. It makes sense later when I join both datasets
daily_pollution_center$city_center <- 1
daily_pollution_out$city_center <- 0
```

```{r}
#Some verifications
df_pollution <- rbind(daily_pollution_center, daily_pollution_out)
unique_values2 <- unique(df_pollution$punto_muestreo)
print(unique_values2)
```

```{r}
# I create a dummy variable that takes the value 1 for dates from 30 November 2018 (date of entry into force of Madrid Central).

df_pollution$post_treat_mad_central<- ifelse(df_pollution$fecha >= as.Date("2018-11-30"), 1, 0)

df_pollution$post_treat_mad_central <- as.integer(df_pollution$post_treat_mad_central)
```


```{r}
# I create a dummy variable that takes the value 1 for dates from the entry into force of Madrid360.
df_pollution$post_treat_mad_360<- ifelse(df_pollution$fecha >= as.Date("2021-12-11"), 1, 0)

df_pollution$post_treat_mad_360<- as.integer(df_pollution$post_treat_mad_360)
```

```{r}

df_pollution$air_quality <- as.numeric(df_pollution$air_quality)

#remove date Na's as the data is 31 days by default even for those months where there are not 31 days.
df_pollution <- df_pollution[!is.na(df_pollution$fecha), ]

df_pollution <- df_pollution |> 
  mutate(punto_muestreo = str_replace(punto_muestreo, "_[^_]+_[^_]+$", ""))

#write.csv(df_pollution, file = "df_pollution.csv", row.names = FALSE)

```

INFORMATION ON CITY CENTRE STATIONS
```{r}
city_stations <- read_delim("informacion_estaciones_red_calidad_aire.csv", delim=";")

city_stations <- city_stations |> 
   select(
     punto_muestreo = CODIGO,
     station_code = CODIGO_CORTO,
     station_name = ESTACION,
     address = DIRECCION,
     station_type = NOM_TIPO,
     longitude = LONGITUD,
     latitude = LATITUD) |> 
    mutate(station_code = as.character(station_code))

#Some modifications to match the names in format with the information of the stations in the peripheral areas so that they can be correctly linked later on.
city_stations <- city_stations |> 
  mutate(
    station_type = str_replace_all(station_type, "tráfico", "trafico"),
    station_type = str_to_title(station_type),
    station_type = str_replace_all(station_type, "\\bfondo\\b", "Fondo"),
    station_type = str_replace_all(station_type, "\\bindustrial\\b", "Industrial"),
    station_type = str_replace_all(station_type, "\\btrafico\\b", "Trafico")
  )
```

INFORMATION ON STATIONS OUTSIDE THE CENTRE OF MADRID

```{r}
com_stations <- read_delim("com_stations.csv", delim=";")

com_stations <- com_stations |> 
  mutate(
    punto_muestreo = estacion_codigo,
    station_code = str_sub(estacion_codigo, -2, -1),
    station_code = if_else(str_starts(station_code, "0"),
                           str_remove(station_code, "^0"),
                           station_code))


  com_stations <- com_stations |> 
  select(station_name = estacion_municipio,
          punto_muestreo,
          station_code,
          station_type = estacion_tipo_estacion,
          area_type = estacion_tipo_area,
          address = estacion_direccion_postal,
          longitude = estacion_coord_longitud,
          latitude = estacion_coord_latitud) 
  
com_stations <- com_stations %>%
  mutate(station_type = paste(area_type, station_type, sep = " ")) |> 
  select(-area_type)


# Fix the latitude and longitude columns using regular expressions
com_stations<- com_stations %>%
  # Remove quotes at the beginning and end of the longitude column
  mutate(longitude = str_remove(longitude, "^\"|\",\\s*\"$")) %>%
  # Replace commas with periods in the longitude column and convert to numeric
  mutate(longitude = as.numeric(gsub(",", ".", longitude))) %>% 
  # Remove quotes and commas from the end of the latitude column
  mutate(latitude = str_remove(latitude, "\"?,?$")) %>%
  # Replace commas with periods in the latitude column and convert to numeric
  mutate(latitude = as.numeric(gsub(",", ".", latitude))) 

# Remove any sequence of two commas AND quotation marks from the address column
com_stations <- com_stations%>% 
  # Remove quotes from the address column
  mutate(address = str_remove(address, '["\']')) %>% 
  # Remove any sequence of two commas from the address column
  mutate(address = str_remove(address, ",{2}"))
```

```{r}
#Check the sampling code of the stations of each dataset.
unique_values1<- unique(daily_pollution_center$punto_muestreo)

unique_values3 <- unique(city_stations$punto_muestreo)

print(unique_values1)
print(unique_values3)

```


```{r}
unique_values2 <- unique(daily_pollution_out$punto_muestreo)
unique_values4 <- unique(com_stations$punto_muestreo)
print(unique_values2) 
print(unique_values4)
```
In the pollution data dataframe there are more stations than we have in the dataframe containing the information of the stations. The missing ones correspond to data from 2023 and have Na's so they are removed to have complete information for all stations.

```{r}
# List of values to remove
valores_a_eliminar <- c("28134002", "28115003", "28106001", "28127004")

# Filter the dataframe to remove observations with those values in punto_muestreo
df_pollution <- df_pollution %>%
  filter(!punto_muestreo %in% valores_a_eliminar)

# Check the changes
print(df_pollution)

```

Join dataframes with information related to pollution stations in the centre of Madrid and peripheral areas.

```{r}
estaciones <- bind_rows(city_stations, com_stations)
```


WEATHER DATA 

```{r}
aemet_api_key("eyJhbGciOiJIUzI1NiJ9.eyJzdWIiOiIxMDA0MDc1NDNAYWx1bW5vcy51YzNtLmVzIiwianRpIjoiMTdlZWU2ZTctY2Y4Yi00N2EzLWE5ZDctYjU1YThmM2Y4OWIyIiwiaXNzIjoiQUVNRVQiLCJpYXQiOjE3MTMwMjM5MjYsInVzZXJJZCI6IjE3ZWVlNmU3LWNmOGItNDdhMy1hOWQ3LWI1NWE4ZjNmODliMiIsInJvbGUiOiIifQ.J2RUdWytG5Y1ZPeHxFZ8pzPeu-QzhHRvHAcOochG0N0", overwrite = TRUE, install = TRUE)
```

```{r}
df_weather_full_2 <- aemet_daily_clim(station = "all", start = "2015-01-01", end = "2023-12-31",                                 verbose = FALSE, return_sf = FALSE) |>  filter(provincia == "MADRID")
```

I recommend to save the file once you have the data because it takes some time...
```{r}
#write.csv(df_weather_full_2, file = "df_weather_full_2.csv", row.names = FALSE)

#df_weather_full_2 <- read_csv("df_weather_full_2.csv")

```

```{r}
weather_stations <- aemet_stations(verbose = FALSE, return_sf = FALSE) %>% 
  filter(provincia == "MADRID") %>% 
  select(-c(indicativo, indsinop, provincia)) %>% 
  rename(altitude = altitud,
         longitude = longitud,
         latitude = latitud)

```

Join climate data with data containing information from stations.
```{r}
weather_data <- inner_join(df_weather_full_2, weather_stations, by = "nombre")
str(weather_data)
```

As the climate stations are not the same as the pollution stations, each pollution station is matched to the nearest climate station. In this way it is possible to have information on temperatures, precipitation, etc. for each pollution station.

```{r}
# Create a matrix of coordinates for air pollution stations
coords_pollution <- as.matrix(estaciones[, c("longitude", "latitude")])

# Create a matrix of coordinates for weather stations
coords_weather <- as.matrix(weather_data[, c("longitude", "latitude")])

# Calculate distances between each air pollution station and each weather station
distances <- distm(coords_pollution, coords_weather, fun = distHaversine)

# Find the index of the closest weather station for each air pollution station
closest_weather_index <- apply(distances, 1, which.min)

# Assign the name of the closest weather station to each air pollution station
estaciones$nombre_meteorologica <- weather_data$nombre[closest_weather_index]

# Check the first rows of the result
head(estaciones)

```

Join the data from each station with the pollution data
```{r}
str(estaciones)
str(df_pollution)

estaciones <- estaciones |> 
  mutate(punto_muestreo = as.character(punto_muestreo))

df_combined <- df_pollution |> 
  left_join(estaciones, by = "punto_muestreo")

#write.csv(df_combined, file = "df_combined.csv", row.names = FALSE)

```
Select the climatic variables of interest
```{r}
weather_data$nombre_meteorologica <- weather_data$nombre
weather_data <- weather_data |> 
  filter(fecha >= as.Date("2017-01-01"))

weather_data <- weather_data |>
  select(c(1, 5:8, 10, 12:14, 16, 18, 20, 21, 23, 25, 29)) |>
  mutate(
    prec = as.numeric(gsub(",", ".", prec)),
    dir = as.numeric(dir),
  )

weather_data <- weather_data |> 
  select(c(1:11, 15:16))
```

For those days for which we do not have weather data, I impute the values.
```{r}
set.seed(123)
m=10
predictor_matrix <- matrix(0, nrow = ncol(weather_data), ncol = ncol(weather_data))
rownames(predictor_matrix) <- colnames(weather_data)
colnames(predictor_matrix) <- colnames(weather_data)

# Especificamos qué variables se utilizarán como predictores para cada variable con valores faltantes
predictor_matrix['tmin', c('fecha', 'tmin')] <- 1
predictor_matrix['tmed', c('fecha', 'tmax', 'tmin')] <- 1
predictor_matrix['tmax', c('fecha', 'tmax', 'tmin')] <- 1
predictor_matrix['dir', c('fecha', 'altitud', 'dir')] <- 1

mice_mod <- mice(weather_data, m=10, predictorMatrix = predictor_matrix)
datos1 <- complete(mice_mod, action=m)


colSums(is.na(datos1))
```


```{r}
#Some modifications before join data
df_combined <- df_combined |> 
  mutate(
    nombre_meteorologica = as.character(nombre_meteorologica),
    fecha = as.Date(fecha)
  )

str(df_combined)

#Data with imputed values
weather_data <- datos1 |> 
  mutate(
    nombre_meteorologica = as.character(nombre_meteorologica),
    fecha = as.Date(fecha)
  )

str(weather_data)

```

Join the dataframes by the column weather_name in df_combined and name in weather_data to have a complete dataframe with, information related to each pollution station, the NO2 pollution values and the weather data. 

```{r}
df_combined_result <- df_combined |> 
  left_join(weather_data, by = c("nombre_meteorologica", "fecha"))
```

```{r}
df_combined_result <- df_combined_result |> 
  arrange(nombre_meteorologica, fecha)

df_combined_imputed <- df_combined_result |> 
  group_by(nombre_meteorologica) |> 
  fill(altitud, tmed, prec, tmin, tmax, dir, velmedia, racha, presMax, presMin, sol, .direction = "down")
```

Save the dataframe to run next script (Models)

```{r}
#write.csv(df_combined_imputed, file = "df_combined_complete.csv", row.names = FALSE)
```
